---
title: "draft package demo"
format: html
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
library(draft)
library(lme4)
library(brms)
set.seed(42)
```

## Simulate data

120 participants nested in 6 clinics. Outcome is a cognitive score (continuous)
and a recovery indicator (binary). Predictors: age, sex, treatment arm.

```{r simulate}
n_sites <- 6
n_per   <- 20
N       <- n_sites * n_per

site        <- rep(paste0("site_", 1:n_sites), each = n_per)
site_effect <- rep(rnorm(n_sites, 0, 0.4), each = n_per)

age      <- round(rnorm(N, mean = 45, sd = 12))
sex      <- factor(sample(c("male", "female"), N, replace = TRUE))
tx       <- factor(sample(c("control", "treatment"), N, replace = TRUE))
tx_num   <- as.integer(tx == "treatment")

# Continuous outcome: cognitive score (0-100)
score <- 55 + 3.5 * tx_num - 0.2 * age +
         2.1 * (sex == "male") + site_effect + rnorm(N, 0, 8)
score <- pmax(0, pmin(100, score))

# Binary outcome: recovery (0/1)
log_odds <- -0.5 + 1.2 * tx_num - 0.02 * age + site_effect
recovery <- rbinom(N, 1, plogis(log_odds))

dat <- data.frame(
  id       = 1:N,
  site     = site,
  age      = age,
  sex      = sex,
  tx       = tx,
  score    = round(score, 1),
  recovery = recovery
)
```

## Table 1

```{r prep-data}
stats <- draft::prep_data(dat)
```

Sample characteristics: the analytic sample included `r stats$age$n` participants
(mean age = `r stats$age$mean`, SD = `r stats$age$sd`). The sample was
`r stats$sex$female$pct`% female and `r stats$tx$treatment$pct`% were randomized
to the treatment arm.

## Linear model: cognitive score

```{r lm}
m_lm <- lm(score ~ tx + age + sex, data = dat)
params_m_lm <- parameters::model_parameters(m_lm)
results_lm <- draft::prep_params(params_m_lm)
```

In the linear model, treatment was associated with a
`r results_lm$txtreatment$coefficient`-point increase in cognitive score
(95% CI: `r results_lm$txtreatment$ci`, `r results_lm$txtreatment$p`).
Each additional year of age was associated with a
`r results_lm$age$coefficient`-point change in score (`r results_lm$age$p`).

## Logistic regression: recovery

```{r glm}
m_glm <- glm(recovery ~ tx + age + sex, data = dat, family = binomial)
params_m_glm <- parameters::model_parameters(m_glm)
results_glm <- draft::prep_params(params_m_glm)
```

In the logistic regression, treatment was associated with greater odds of
recovery (log-OR = `r results_glm$txtreatment$coefficient`,
95% CI: `r results_glm$txtreatment$ci`, `r results_glm$txtreatment$p`).

## Mixed model: score with site-level random intercepts

```{r glmer}
m_glmer <- lmer(score ~ tx + age + sex + (1 | site), data = dat)
params_m_glmer <- parameters::model_parameters(m_glmer, effects = "fixed")
results_glmer <- draft::prep_params(params_m_glmer)
```

Accounting for clustering within sites, the treatment effect on score was
`r results_glmer$txtreatment$coefficient` (95% CI: `r results_glmer$txtreatment$ci`,
`r results_glmer$txtreatment$p`).

## Bayesian mixed model: recovery with site-level random intercepts

Fast sampling settings for demo purposes (increase iter for real analyses).

```{r brm, results='hide'}
m_brm <- brm(
  recovery ~ tx + age + sex + (1 | site),
  data   = dat,
  family = bernoulli(),
  chains = 2,
  iter   = 1000,
  warmup = 400,
  cores  = 2,
  seed   = 42,
  file = "m1.rds",
  file_refit = "on_change",
  silent = 2
)
params_m_brm <- parameters::model_parameters(m_brm, effects = "all")
results_brm <- draft::prep_params(params_m_brm)
```


```{r}
mycorr = correlation::correlation(data.frame(x = rnorm(10, 0, 1), y = rnorm(5, 5, 1)))
```

In the Bayesian model, the posterior median log-OR for treatment was
`r results_brm$txtreatment$median`
(95% CrI: `r results_brm$txtreatment$ci`,
pd = `r results_brm$txtreatment$pd`).

---

## Launch draft app

Run this interactively (not during render) to open the app with all
objects above available for browsing:


```{r launch, eval=FALSE}
draft::launch_app()
```

