---
title: "dev — modelbased objects"
format: html
---

## Setup

```{r setup}
library(modelbased)
library(lme4)
library(draft)
```

## Models

```{r models}
# Simple linear model — continuous predictors + moderator
m_lm <- lm(mpg ~ wt + cyl + am, data = mtcars)

# Model with a factor — needed for meaningful means and contrasts
m_factor <- lm(mpg ~ factor(cyl) + wt, data = mtcars)

# Mixed model — needed for estimate_grouplevel()
m_lmer <- lmer(Reaction ~ Days + (Days | Subject), data = sleepstudy)
```

## estimate_means()

Marginal means at each level of a factor predictor, averaging over other predictors.

```{r estimate-means}
means_cyl <- estimate_means(m_factor, by = "cyl")
means_cyl

class(means_cyl)
names(means_cyl)
```

## estimate_contrasts()

Pairwise differences between factor levels with CIs and p-values.

```{r estimate-contrasts}
contrasts_cyl <- estimate_contrasts(m_factor, contrast = "cyl")
contrasts_cyl

class(contrasts_cyl)
names(contrasts_cyl)
```

## estimate_slopes()

Marginal effect (slope) of a continuous predictor, optionally at levels of a moderator.

```{r estimate-slopes}
# Overall slope
slopes_wt <- estimate_slopes(m_lm, trend = "wt")
slopes_wt

# Slope at each level of a moderator
slopes_wt_by_am <- estimate_slopes(m_lm, trend = "wt", by = "am")
slopes_wt_by_am

class(slopes_wt)
names(slopes_wt)
names(slopes_wt_by_am)
```

## estimate_grouplevel()

Random effect estimates per group from a mixed model.

```{r estimate-grouplevel}
group_fx <- estimate_grouplevel(m_lmer)
group_fx

class(group_fx)
names(group_fx)
```

## What does draft see?

```{r for-draft}
# Column names across all four — drives chip generation in the inspector
names(means_cyl)
names(contrasts_cyl)
names(slopes_wt)
names(group_fx)

# Are classes consistent enough to share a single inspector renderer?
c(
  means     = class(means_cyl)[1],
  contrasts = class(contrasts_cyl)[1],
  slopes    = class(slopes_wt)[1],
  grouplevel = class(group_fx)[1]
)
```

## draft::launch_app() — manual test

Uncomment to launch the app and inspect these objects.

```{r launch, eval=FALSE}
draft::launch_app()
```
