---
title: "dev — effectsize objects"
format: html
---

## Setup

```{r setup}
library(effectsize)
library(draft)
```

## Data

```{r data}
# Continuous outcome by group — for difference measures
# iris for multi-group ANOVA
# mtcars for binary group (am) and contingency tables
```

## Difference measures — effectsize_difference

Single-row output. Flat inline paths: `es_d_$cohens_d`, `es_d_$ci_low`, `es_d_$ci_high`.

```{r difference}
# Cohen's d — continuous outcome by binary group
es_d  <- cohens_d(mpg ~ am, data = mtcars)

# Hedges' g — bias-corrected version of Cohen's d
es_g  <- hedges_g(mpg ~ am, data = mtcars)

# Glass's delta — uses only control group SD
es_gl <- glass_delta(mpg ~ am, data = mtcars)

# Rank-biserial — non-parametric equivalent
es_rb <- rank_biserial(mpg ~ am, data = mtcars)

es_d
es_g
```

## ANOVA effect sizes — effectsize_anova

Multi-row output keyed by predictor. Nested inline paths:
`es_eta_$cyl$eta2_partial`, `es_eta_$cyl$ci_low`.

```{r anova}
m_aov <- aov(mpg ~ cyl + gear + am, data = mtcars)

# Partial eta-squared
es_eta   <- eta_squared(m_aov, partial = TRUE)

# Omega-squared — less biased than eta-squared
es_omega <- omega_squared(m_aov)

# Epsilon-squared
es_eps   <- epsilon_squared(m_aov)

es_eta
es_omega
```

## Association measures — effectsize_table

Single-row output. Flat inline paths: `es_v_$cramers_v_adjusted`, `es_v_$ci_low`.

```{r association}
# Cramér's V — strength of association in a contingency table
es_v   <- cramers_v(table(mtcars$cyl, mtcars$gear))

# Cohen's w — similar to Cramér's V
es_w   <- cohens_w(table(mtcars$cyl, mtcars$am))

# Phi — 2x2 tables only
es_phi <- phi(table(mtcars$am, mtcars$vs))

es_v
es_w
```

## Launch app

```{r launch}
#| eval: false
draft::launch_app()
```
