---
title: "Demo Analysis — Sleep, Cognition, and Anxiety"
subtitle: "A simulated longitudinal study"
format: html
---

## Overview

This is a simulated study examining how sleep quality and anxiety predict
cognitive performance, with participants measured across two time points.
The script is designed to populate the draft environment with a variety of
object types for testing the app.

## Setup

```{r setup}
#| message: false
library(lme4)
library(parameters)
library(modelbased)
library(performance)
library(effectsize)
library(correlation)
library(draft)

set.seed(42)
```

## Simulate data

```{r simulate}
n_subjects  <- 80
n_timepoints <- 2

subject_id  <- rep(1:n_subjects, each = n_timepoints)
time        <- rep(c(0, 1), times = n_subjects)

# Between-person variables
anxiety     <- rep(rnorm(n_subjects, mean = 10, sd = 3), each = n_timepoints)
age         <- rep(round(runif(n_subjects, 20, 65)), each = n_timepoints)
condition   <- rep(sample(c("control", "treatment"), n_subjects, replace = TRUE),
                   each = n_timepoints)

# Sleep quality (0-10; higher = better); improves slightly with treatment over time
sleep <- 5 + 0.8 * (condition == "treatment") * time +
         rnorm(n_subjects * n_timepoints, sd = 1.2)
sleep <- pmax(0, pmin(10, sleep))

# Cognitive score: driven by sleep, age, anxiety, and random subject intercept
subject_intercept <- rep(rnorm(n_subjects, sd = 3), each = n_timepoints)
cognition <- 60 + 2.5 * sleep - 0.15 * age - 0.8 * anxiety +
             subject_intercept + rnorm(n_subjects * n_timepoints, sd = 4)

# Binary outcome: high performer (cognition above median)
high_performer <- as.integer(cognition > median(cognition))

df <- data.frame(
  subject_id     = factor(subject_id),
  time           = time,
  condition      = factor(condition, levels = c("control", "treatment")),
  age            = age,
  anxiety        = anxiety,
  sleep          = sleep,
  cognition      = cognition,
  high_performer = high_performer
)

head(df)
```

## Descriptives

```{r descriptives}
# Summary of continuous variables at baseline
baseline <- df[df$time == 0, ]
desc     <- baseline[, c("age", "anxiety", "sleep", "cognition")]
```

## Distribution summaries — datawizard

```{r dist-unstratified}
# Unstratified: flat list keyed by variable name
# Paths: dist_baseline_$age$mean, dist_baseline_$cognition$sd
library(datawizard)
dist_baseline <- describe_distribution(desc)
dist_baseline
```

```{r dist-stratified}
# Stratified by condition: nested list keyed by group then variable
# Paths: dist_by_condition_$control$age$mean, dist_by_condition_$treatment$cognition$sd
dist_by_condition <- describe_distribution(
  baseline[, c("age", "anxiety", "sleep", "cognition", "condition")],
  by = "condition"
)
dist_by_condition
```

## Correlation matrix

```{r correlations}
# Pairwise correlations among continuous predictors and outcome (baseline only)
corr_matrix <- correlation(desc)
corr_matrix
```
```{r}
corr_matrix_ <- draft::prep_params(corr_matrix)
```

                
              

## Linear model — predictors of cognition (cross-sectional, baseline only)

```{r lm}
m_lm <- lm(cognition ~ sleep + anxiety + age + condition, data = baseline)

params_lm <- model_parameters(m_lm)
params_lm
```

## Linear model — performance

```{r lm-performance}
perf_lm <- model_performance(m_lm)
perf_lm
```

## Effect sizes — linear model predictors

```{r lm-effectsize}
# Standardized coefficients as effect sizes
es_std <- effectsize(m_lm)
es_std
```

## Marginal means — condition effect on cognition

```{r marginal-means}
# Estimated marginal means of cognition by condition, holding other vars at mean
means_condition <- estimate_means(m_lm, at = "condition")
means_condition
```

## Contrasts — treatment vs control

```{r contrasts}
contrasts_condition <- estimate_contrasts(m_lm, contrast = "condition")
contrasts_condition
```

## Mixed model — longitudinal (all time points)

```{r glmer}
# Binary outcome: probability of being a high performer
# Fixed effects: sleep, anxiety, time, condition
# Random intercept by subject
m_glmer <- glmer(
  high_performer ~ sleep + anxiety + time * condition + (1 | subject_id),
  data   = df,
  family = binomial
)

params_glmer <- model_parameters(m_glmer)
params_glmer
```

## Mixed model — performance

```{r glmer-performance}
perf_glmer <- model_performance(m_glmer)
perf_glmer
```

## Model comparison

```{r compare}
# Null model for comparison
m_glmer_null <- glmer(
  high_performer ~ time + (1 | subject_id),
  data   = df,
  family = binomial
)

comparison <- compare_performance(m_glmer_null, m_glmer, rank = TRUE)
comparison
```

## Marginal means from mixed model — condition x time

```{r mixed-means}
means_time_condition <- estimate_means(m_glmer, at = c("condition", "time"))
means_time_condition
```

## Marginal slopes — effect of sleep over time

```{r slopes}
slopes_sleep <- estimate_slopes(m_glmer, trend = "sleep", at = "time")
slopes_sleep
```

## Cohen's d — cognition by condition (baseline)

```{r cohens-d}
es_d <- cohens_d(cognition ~ condition, data = baseline)
es_d
```

## Eta-squared — ANOVA decomposition

```{r eta}
m_aov <- aov(cognition ~ condition + cut(age, 3) + cut(anxiety, 3), data = baseline)
es_eta <- eta_squared(m_aov)
es_eta
```

## Effect size conversions — plain scalars

These return plain numeric values (no special class), so they appear as scalar
objects in the env panel rather than going through the effectsize handler.

```{r conversions}
# Convert Cohen's d to Pearson r
d_as_r <- d_to_r(es_d$Cohens_d)

d_as_r
```

## Launch the app

```{r launch}
#| eval: false
draft::launch_app()
```
