---
title: "dev — datawizard describe_distribution objects"
format: html
---

## Setup

```{r setup}
#| message: false
library(datawizard)
library(draft)
```

## Data

Using `iris` (dot-separated column names) and `mtcars` (no dots) to show how
`sanitize_key()` handles different naming conventions.

Note: column names with dots (`Sepal.Length`) become `sepallength` (dots stripped).
Column names with underscores or spaces produce cleaner keys.

## Unstratified — describe_distribution()

Single flat list keyed by variable name.
Paths: `dist_iris_$sepallength$mean`, `dist_iris_$sepallength$sd`

```{r flat-iris}
dist_iris <- describe_distribution(iris[, 1:4])
dist_iris
```

```{r flat-mtcars}
# mtcars column names have no dots — keys are cleaner
dist_mtcars <- describe_distribution(mtcars[, c("mpg", "wt", "hp", "disp")])
dist_mtcars
# Paths: dist_mtcars_$mpg$mean, dist_mtcars_$wt$sd
```

## Stratified — describe_distribution(by = )

Nested list: group → variable → statistic.
Paths: `dist_by_species_$setosa$sepallength$mean`

```{r stratified-species}
dist_by_species <- describe_distribution(iris, by = "Species")
dist_by_species
```

```{r stratified-cyl}
# mtcars stratified by number of cylinders
dist_by_cyl <- describe_distribution(
  mtcars[, c("mpg", "wt", "hp", "cyl")],
  by = "cyl"
)
dist_by_cyl
# Paths: dist_by_cyl_$`4`$mpg$mean, dist_by_cyl_$`6`$wt$sd
# Note: numeric group keys (4, 6, 8) need backtick-quoting in $ paths
```

## Multiple by variables

```{r stratified-multi}
# Two grouping variables — keys are pasted together
mtcars$am_label  <- ifelse(mtcars$am == 1, "manual", "automatic")
mtcars$cyl_label <- paste0("cyl", mtcars$cyl)

dist_multi <- describe_distribution(
  mtcars[, c("mpg", "wt", "am_label", "cyl_label")],
  by = c("am_label", "cyl_label")
)
dist_multi
```

## Launch the app

```{r launch}
#| eval: false
draft::launch_app()
```
