---
title: "Getting Started with draft"
format: html
execute:
  eval: false
  echo: true
---

## Overview

Writing a results section means translating your R objects into prose,
keeping that prose accurate any time your data or analysis updates, and avoiding
errors in transcribing values from your data into prose. `draft` is a
local Shiny app that makes this process systematic. It reads objects from your
R environment, structures their values into stable named lists, and gives you a
live text editor where inline references resolve as you type.

When you're done writing, you copy two things into your `.qmd` or `.Rmd`
document: a small setup chunk that rebuilds the lists, and inline references
like `params_m1_$wt$coefficient` that render correctly every time you knit.

`draft` leverages the `easystats` suite of packages (https://easystats.github.io/easystats/) which
have a systematic and consistent structure, relatively few dependencies, and coverage for many 
different model types and analytical approaches in R. `draft` also supports reporting of bespoke
dataframes you create or other values in your R environment. 

## Installation

```{r install, eval = FALSE}
# Install from GitHub
remotes::install_github("rbcavanaugh/draft")
```

`draft` works best alongside the
[easystats](https://easystats.github.io/easystats/) packages. Install the ones
you use:

```{r install-deps, eval = FALSE}
install.packages(c("parameters", "effectsize", "modelbased", "performance", "correlation"))
```

## A simple workflow

**Step 1: Run your analysis as normal**

Fit your models and prepare your objects in R. Nothing special required.

```{r basic, eval = FALSE}
library(draft)
library(parameters)

m1        <- lm(mpg ~ wt + cyl, data = mtcars)
params_m1 <- model_parameters(m1)
```

**Step 2: Launch the app**

```{r launch, eval = FALSE}
launch_app()
```

The app reads everything in your current environment automatically.

**Step 3: Browse your objects**

In the left panel, click `params_m1`. The inspector shows each predictor as a
row of values (estimate, SE, CI, p). Each shows the value and the path you'd
use to reference it inline.

**Step 4: Write prose in the editor**

In the right panel, type your results sentence. Use `{...}` to insert inline
references — click the value you want to reference and paste the path in:

```
Weight was negatively associated with fuel economy
(b = {params_m1_$wt$coefficient}, 95% CI [{params_m1_$wt$ci_low},
{params_m1_$wt$ci_high}], p = {params_m1_$wt$p}).
```

The preview below the editor renders the actual values as you type, so you can
see exactly what will appear in your document.

**Step 5: Copy the inline code**

Click **Copy inline code**. The `{...}` delimiters are converted to backtick-r
inline expressions — paste the result directly into your `.qmd` or `.Rmd`:

```{r show-rmd-output, eval=TRUE, echo=FALSE, results='asis'}
cat("```\nWeight was negatively associated with fuel economy\n",
    "(b = `r params_m1_$wt$coefficient`, 95% CI [`r params_m1_$wt$ci_low`,\n",
    "`r params_m1_$wt$ci_high`], p = `r params_m1_$wt$p`).\n```",
    sep = "")
```

**Step 6: Copy the setup chunk**

Click **Copy chunk**. This gives you the setup code that rebuilds the list —
paste it into your document's setup chunk:

```{r setup-chunk, eval = FALSE}
params_m1_ <- draft::prep_params(params_m1)
```

**Step 7: Re-render your document**

Knit or render your `.qmd`/`.Rmd` from scratch to confirm the values resolve
correctly and match what you saw in the app.

> The app never writes to your document. You paste two things: the setup chunk
> (once, in your setup block) and the inline references (wherever you report
> results). Everything else stays in your script.

---

## Working with `parameters`

`model_parameters()` is the main entry point for regression models. It works
consistently across `lm`, `glm`, `lmer`, `glmer`, `brms`, and many others —
the output always has the same structure, and `draft` handles all of them the
same way.

```{r params, eval = FALSE}
library(parameters)

# Linear model
params_lm   <- model_parameters(lm(mpg ~ wt + cyl, data = mtcars))

# Logistic regression
params_glm  <- model_parameters(glm(am ~ wt + hp, data = mtcars, family = binomial))

# Mixed model
library(lme4)
params_lmer <- model_parameters(lmer(Reaction ~ Days + (Days | Subject),
                                     data = sleepstudy), effects = "fixed")

# Bayesian — same workflow, different field names (median, pd instead of coefficient, p)
# params_bayes <- model_parameters(brm(...))
```

The inline path pattern is always `list_name_$term$field`. Field names vary by
model type — frequentist models produce `coefficient` and `p`; Bayesian models
produce `median` and `pd`. **The inspector always shows you the exact field
names for your specific model** — check there rather than assuming.

Correlation matrices from the `correlation` package work too, since
`correlation()` returns a `parameters_model` object:

```{r correlation, eval = FALSE}
library(correlation)
corr_matrix <- correlation(mtcars[, 1:4])
# e.g. corr_matrix_$mpg_wt$r, corr_matrix_$mpg_wt$p
```

**The app doesn't control your statistics.** Confidence levels, standardization,
and any other options stay in your script. To use 89% CIs instead of 95%:

```{r params-options, eval = FALSE}
params_lm_89  <- model_parameters(m1, ci = 0.89)
params_lm_std <- model_parameters(m1, standardize = "refit")
```

Relaunch the app after changing your script to see updated values.

---

## Working with `effectsize`

Effect size objects come in two shapes depending on the function.

**Single-effect functions** (e.g. `cohens_d()`, `cramers_v()`) return a
single-row result that preps into a **flat** named list — the path goes directly
to the field:

```{r es-single, eval = FALSE}
library(effectsize)

es_d <- cohens_d(mpg ~ am, data = mtcars)
# es_d_$cohens_d, es_d_$ci_low, es_d_$ci_high

es_v <- cramers_v(table(mtcars$cyl, mtcars$gear))
# es_v_$cramers_v_adjusted, es_v_$ci_low
```

**Multi-effect functions** (e.g. `eta_squared()` with several predictors) return
one row per predictor and prep into a **nested** list keyed by predictor name:

```{r es-multi, eval = FALSE}
es_eta <- eta_squared(aov(mpg ~ cyl + gear, data = mtcars))
# es_eta_$cyl$eta2_partial, es_eta_$cyl$ci_low
# es_eta_$gear$eta2_partial
```

Conversion functions like `d_to_r()` return plain numeric scalars rather than
effectsize objects. They appear in the environment panel as scalar values (see
[Working with scalars] below).

---

## Working with `modelbased`

`modelbased` produces marginal means, contrasts, and slopes — each prepped into
a named list keyed by the relevant level or comparison.

```{r modelbased, eval = FALSE}
library(modelbased)

m <- lm(mpg ~ factor(cyl) + wt, data = mtcars)

# Marginal means — one entry per level
means_cyl     <- estimate_means(m, at = "cyl")
# means_cyl_$`4`$mean, means_cyl_$`6`$mean

# Contrasts — keys use the _vs_ pattern
contrasts_cyl <- estimate_contrasts(m, contrast = "cyl")
# contrasts_cyl_$`4_vs_6`$difference, contrasts_cyl_$`4_vs_6`$ci_low

# Marginal slopes
m_int      <- lm(mpg ~ wt * factor(cyl), data = mtcars)
slopes_wt  <- estimate_slopes(m_int, trend = "wt", at = "cyl")
# slopes_wt_$`4`$coefficient, slopes_wt_$`6`$coefficient
```

Keys come directly from your factor level names. Descriptive levels like
`"control"` and `"treatment"` produce readable paths; numeric levels like `4`,
`6`, `8` need backtick-quoting in `$` paths — the inspector handles this
automatically when you click to copy a value.

---

## Working with `performance`

`model_performance()` returns a flat list of fit indices for a single model.
`compare_performance()` returns a nested list keyed by model name.

```{r performance, eval = FALSE}
library(performance)

perf_lm    <- model_performance(m1)
# perf_lm_$r2, perf_lm_$aic, perf_lm_$rmse

comparison <- compare_performance(
  m_null <- lm(mpg ~ 1, data = mtcars),
  m_full <- lm(mpg ~ wt + cyl + am, data = mtcars)
)
# comparison_$m_full$aic, comparison_$m_full$r2
```

Available indices depend on the model type — mixed models include ICC, logistic
models include Tjur's R². The inspector shows exactly what your model provides.

---

## Working with `datawizard`

`datawizard::describe_distribution()` produces a distribution summary table.
`draft` supports both the plain (unstratified) and stratified (`by =`) forms.

**Unstratified** — one entry per variable, flat list:

```{r dist-flat, eval = FALSE}
library(datawizard)

dist_baseline <- describe_distribution(df[, c("age", "anxiety", "sleep")])
# dist_baseline_$age$mean, dist_baseline_$age$sd
# dist_baseline_$sleep$median
```

**Stratified** — the most common use: one entry per group, then per variable:

```{r dist-stratified, eval = FALSE}
dist_by_condition <- describe_distribution(
  df[, c("age", "sleep", "condition")],
  by = "condition"
)
# dist_by_condition_$control$age$mean
# dist_by_condition_$treatment$sleep$sd
```

Multiple `by` variables are supported; group keys are pasted together with `_`.
Note that column names with dots (`Sepal.Length`) become compacted keys
(`sepallength`) — check the inspector for the exact key names.

---

## Working with data frames

Any `data.frame` or tibble in your environment appears in the environment panel.
`draft` summarises each column automatically:

- **Continuous columns** (numeric with more than 10 unique values) — `mean`,
  `sd`, `median`, `min`, `max`, `n`, `n_missing`
- **Categorical columns** (factor, character, or numeric with ≤ 10 unique
  values) — `n` and `n_missing` overall, plus `n` and `pct` per level
- **High-cardinality factor/character columns** (e.g. subject IDs, free-text)
  — excluded; no meaningful summary to copy

```{r dataframe, eval = FALSE}
# prep_data() is what the app calls — you can also call it directly
desc_ <- draft::prep_data(mtcars)

# desc_$mpg$mean, desc_$mpg$sd
# desc_$cyl$`4`$n, desc_$cyl$`4`$pct
```

The threshold between continuous and categorical defaults to 10 unique values
and can be adjusted:

```{r cat-threshold, eval = FALSE}
desc_ <- draft::prep_data(my_df, cat_threshold = 20)
```

---

## Working with scalars

Any single value — numeric, character, or logical — assigned to a variable
appears in the environment panel as a scalar. No setup code is needed. Reference
it directly by name in your prose:

```{r scalars, eval = FALSE}
n_participants <- 120
study_site     <- "Pittsburgh"

# {n_participants} participants were recruited from {study_site}.
# → 120 participants were recruited from Pittsburgh.
```

Numeric scalars are displayed rounded to two decimal places. Conversion
functions from `effectsize` land here:

```{r scalar-conversions, eval = FALSE}
es_d   <- cohens_d(mpg ~ am, data = mtcars)
d_as_r <- d_to_r(es_d$Cohens_d)
# d_as_r appears as a scalar ≈ 0.24
```

---

## Tips

- **Everything inside `{}` is evaluated as R.** You are not limited to plain
  variable names — any valid R expression works. To report a proportion as a
  percentage, write `{my_value * 100}%` in the editor. Other examples:
  `{abs(b)}` to drop a sign, `{n - 1}` for degrees of freedom,
  `{ifelse(p < .05, "significant", "non-significant")}` for conditional language.

- **The app has no statistical settings of its own.** All decisions about what
  to report — confidence level, effect size type, standardization — live in your
  script as function arguments. Change them there and relaunch.

- **Check field names in the inspector.** Field names vary by model type and
  package version. The inspector shows the exact names for your specific object;
  always check there rather than relying on documentation examples.

- **The trailing underscore is intentional.** The prep list for `params_m1` is
  stored as `params_m1_`. The underscore distinguishes it from the original
  model object in your prose and prevents name collisions.

- **Numeric formatting follows APA style automatically.** Non-p values are
  rounded to two decimal places and scientific notation is suppressed. P-values
  are formatted to three decimal places with `< .001` below the threshold.
  You do not need to round values yourself in the editor.
